## Statistical Inference - Quiz 2

### Question 1

What is the variance of the distribution of the average an IID draw of n observations from a population with mean $μ$ and variance $σ^2$?

-   $\frac{σ^2}{n}$
-   $σ^2$
-   $2\frac{σ}{\sqrt{n}}$
-   $\frac{σ}{n}$

Answer : $\frac{σ^2}{n}$

### Question 2

Suppose that diastolic blood pressures (DBPs) from men aged 35-44 are normally distributed with a mean of 80mmHg and a standard deviation of 10 mmHg. About what is the probability that a random 35-44 year old has a DBP less than 70?

-   22%
-   32%
-   8%
-   16%

```{r q2}
pnorm(70,mean=80,sd=10,lower.tail = TRUE)
```

Answer : The searched probability is around 16%.

### Question 3

Brain volume for adult women is normally distributed with a mean of about 1,100 cc for women with a standard deviation of 75 cc. What brain volume represents the 95th percentile?

-   approximately 1247
-   approximately 1223
-   approximately 1175
-   approximately 977

```{r q3}
qnorm(0.95,mean=1100,sd=75,lower.tail = TRUE)
```

Answer : The 95th quantile for this distribution is approximately 1223.

### Question 4

Refer to the previous question. Brain volume for adult women is normally distributed with a mean of about 1,100 cc for women with a standard deviation of 75 cc. Consider the sample mean of 100 random adult women from this population. What is th 95th percentile of the distribution of the sample mean?

-   approximately 1112
-   approximately 1088
-   approximately 1110
-   approximately 1115

```{r q4}
qnorm(0.95,mean=1100,sd=75/10,lower.tail = TRUE)
```

Answer : The 95th quantile for this distribution mean is approximately 1112.

### Question 5

You flip a fair coin 5 times. About what is the probability of getting 4 or 5 heads?

-   6%
-   19%
-   3%
-   12%

As the coin is fair, the probability of getting 1 head at each flip is 0.5. The probability of getting at least 4 heads after 5 flips can be computed using the binomial law : $p=C_4^5(0.5)^4(1−0.5)+C_5^5(0.5)^5$

```{r q5}
pbinom(3, size=5, prob=0.5, lower.tail = FALSE)
```

Answer : The probability is approximately 19%.

### Question 6

The respiratory disturbance index (RDI), a measure of sleep disturbance, for a specific population has a mean of 15 (sleep events per hour) and a standard deviation of 10. They are not normally distributed. Give your best estimate of the probability that a sample mean RDI of 100 people is between 14 and 16 events per hour.

-   68%
-   95%
-   34%
-   47.5%

Answer : The standard error of the mean is $\frac{σ}{\sqrt{n}}$ , where $σ=10$ and $n=100$. So the standard error is 1. The probability of the RDI being inside 1 standard deviation around the mean is 68%.

### Question 7

Consider a standard uniform density. The mean for this density is 0.5 and the variance is 1/12. You sample 1000 observations from this distribution and take the sample mean, what value would you expect it to be near?

-   0.75
-   0.25
-   0.10
-   0.5

Answer : Using the LLN, the value should be approximately 0.5.

### Question 8

The number of people showing up at a bus stop is assumed to be Poisson with a mean of 5 people per hour. You watch the bus stop for 3 hours. About what's the probability of viewing 10 of fewer people?

-   0.08
-   0.12
-   0.03
-   0.06

```{r q8}
ppois(10, lambda=15)
```

Answer : The probability is approximately 12%.
