The most harmful with respect to population health and the greatest economic consequences events.
============================================================================

## Synopsis
This project involves exploring the U.S. National Oceanic and Atmospheric Administration's (NOAA) storm database and answers some questions about severe weather events. The events in the database start in the year 1950 and end in November 2011 and includes fatalities, injuries, property and crop damage.  Tornado is the most harmful type of events with respect to population health. 7 types of fatality events and 4 types of injury events are at least 5% of the number of tornado events. Flood and drought are the most harmful type of events with respect to property and crop damage respectively. 7 types of property damage and 10 types of crop damage are at least 5% of the property damage of flood and crop damage of drought respectively.

## 1.Data Processing
### Downloading the dataset:
```{r download-data, cache=TRUE}
filename <- "repdata_data_StormData.csv.bz2"
if (!file.exists(filename)){
        fileURL <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2"
        download.file(fileURL, filename, method="curl")
}  
data <- read.csv(filename)
```

### Selecting and preprocessing variables
Select variables for analysis and convert alphabetical characters used to signify magnitude (“K” for thousands, “M” for millions, and “B” for billions) to a numeric ones.
```{r processing, cache=TRUE}
DataDamage <- select(data, EVTYPE, FATALITIES, INJURIES, PROPDMG, PROPDMGEXP, CROPDMG, CROPDMGEXP)
DataDamage$PROPDMGEXP <- as.numeric(factor(DataDamage$PROPDMGEXP, levels = c('K', 'M', 'B'), labels = c(1, 2, 3)))
DataDamage$CROPDMGEXP <- as.numeric(factor(DataDamage$CROPDMGEXP, levels = c('K', 'M', 'B'), labels = c(1, 2, 3)))
```

### Library loading
```{r dplyr}
library(dplyr)
```

## 2.Results
### Fatality and injurity events
Group the dataset by type of event (EVTYPE) and find the sums of fatalities and injuries for every event, then find the event with maximum numbers of fatalities and injuries
```{r HealthDamage, cache=TRUE}
HealthDamage <- group_by(DataDamage, EVTYPE) %>% 
                summarise(Fatalities = sum(FATALITIES, na.rm = TRUE), 
                          Injuries = sum(INJURIES, na.rm = TRUE))
MaxFatalEvent <- HealthDamage$EVTYPE[which.max(HealthDamage$Fatalities)]
NumFataEvent <- max(HealthDamage$Fatalities)
MaxInjuryEvent <- HealthDamage$EVTYPE[which.max(HealthDamage$Injuries)]
NumInjuryaEvent <- max(HealthDamage$Injuries)
```
Event with maximum numbers of fatalities: `r MaxFatalEvent` with numbers of cases `r NumFataEvent`  
Event with maximum numbers of injuries: `r MaxInjuryEvent` with numbers of cases `r NumInjuryaEvent` 

Make boxplots of values Fataliteis and Injuries for event types
```{r graph_HealthDamage}
par(mfrow = c(1,2), mar = c(4, 4, 2, 1), oma = c(0, 0, 2, 0))
with(HealthDamage, {
boxplot(Fatalities, xlab = 'Fatalities', ylab = 'The number of occurrences')
boxplot(Injuries, xlab = 'Injuries', ylab = 'The number of occurrences')
mtext('Health Damage', outer = TRUE)
})
```

We see that some other types of events also have significant numbers of occurrences. So we choose events whose number of fatalities is at least 5% of the number of occurrences of `r MaxFatalEvent` and number of injuries is at least 5% of the number of occurrences of `r MaxInjuryEvent`
```{r table_HealthDamage}
print(select(HealthDamage, EVTYPE, Fatalities) %>% filter(Fatalities > 0.05 * NumFataEvent))
print(select(HealthDamage, EVTYPE, Injuries) %>% filter(Injuries > 0.05 * NumInjuryaEvent))
```

### Property and crop damage events
Group the dataset by EVTYPE and find the sum of property and crop damages for every event, then find the event with maximum value of property and crop damages
```{r EconomicDamage, cache=TRUE}
EconomicDamage <- group_by(DataDamage, EVTYPE) %>% 
                summarise(PropDmg = sum(PROPDMG*1000^PROPDMGEXP, na.rm = TRUE), 
                          CropDmg = sum(CROPDMG*1000^CROPDMGEXP, na.rm = TRUE))
MaxPropDmgEvent <- EconomicDamage$EVTYPE[which.max(EconomicDamage$PropDmg)]
PropDmgEvent <- max(EconomicDamage$PropDmg)
MaxCropDmgEvent <- EconomicDamage$EVTYPE[which.max(EconomicDamage$CropDmg)]
CropDmgEvent <- max(EconomicDamage$CropDmg)
```
Event with maximum of property damages: `r MaxPropDmgEvent` with damage `r PropDmgEvent` dollars  
Event with maximum of crop damages: `r MaxCropDmgEvent` with damage `r CropDmgEvent` dollars  
Make boxplots of values property and crop damages for event types
```{r graph_EconomicDamage}
par(mfrow = c(1,2), mar = c(4, 4, 2, 1), oma = c(0, 0, 2, 0))
with(EconomicDamage, {
boxplot(PropDmg, xlab = 'Property Damage', ylab = 'Dollars')
boxplot(CropDmg, xlab = 'Crop Damage', ylab = 'Dollars')
mtext('Economic Damage', outer = TRUE)
})
```

We see that some other types of events also have significant economic damages. So we choose events whose property damages are at least 5% of the damage of `r MaxPropDmgEvent` and crop damages are at least 5% of the of the damage of `r MaxCropDmgEvent`
```{r table_EconomicDamage}
print(select(EconomicDamage, EVTYPE, PropDmg) %>% filter(PropDmg > 0.05 * PropDmgEvent))
print(select(EconomicDamage, EVTYPE, CropDmg) %>% filter(CropDmg > 0.05 * CropDmgEvent))
```
